\section{Introduction and motivation}
\label{sec:intr-motiv}

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for correct by construction code generation
and formally verifying functional properties of developed programs, but
leave nonfunctional properties such as timing behavior as an
implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the speed of computation of the program for
different input signals. Unlike a discrete physical clock, which has
fixed real-time period, the period of the logical clock is elastic.
Although the timing model with logical clock has worked well for
reactive synchronous/GALS languages in designing discrete systems, there
is a need to introduce real-time since often, implementation models are
in the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}. This paper targets
enhancing formal GALS and synchronous languages, by adding real-time
constructs. The real-time constructs speed up a design process by
allowing system designers to explicitly program real-time postponements
within the language itself without need of external resources such as
timers.

%Moreover, adding timing
%capabilities as first class language constructs create additional
%opportunity for formally verifying functional and non-functional
%requirements before system deployment.

\input{motivation}

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
\hj{(1) real-time wait does not affect the model of logical time in
	SystemJ program as long as the amount of delay is within boundaries
that can be determined statically by program analysis.} (2) the waiting
model relies on the use of relative instead of absolute real-time, i.e.,
a wait in selected time units is counted from the beginning of the wait
statement and (3) these waiting mechanisms allow mixing behaviors with
real-time features and others with only logical time. All these as well
as semantics of the wait constructs are discussed and illustrated in the
rest of the paper. Our contributions can be further refined as follows:

\begin{enumerate*}
\item Real-time is specified in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
% \item We consider the property of \textit{non
%     maximal-parallelism}. \textit{Non-maximal parallelism:} There are
%   not always sufficient resources for processes to execute, so
%   effectively processes are allocated and scheduled with different
%   optimization criteria -- reducing computation time, reducing power
%   consumption, etc. This allocation and scheduling is an integral part
%   of any language compilation and needs to be considered when
%   introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We shed light upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\item We \textbf{assume} that real-time analyzable platforms are
	available on which SystemJ programs are executed.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
in the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:proof-correctness} sketches the correctness
proofs for the introduced
algorithms. Section~\ref{sec:experimental-results} gives the
experimental results for a set of real-time applications. In
Section~\ref{sec:disc-perc-limit} we discuss the advantages and
limitations of the proposed approach. Section~\ref{sec:related-work}
positions this work in relation to other approaches to introducing
real-time in other languages. Finally, we conclude in
Section~\ref{sec:concl-future-work}.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: 
